Combinatorial Proofs and Decomposition Theorems for First-order Logic

### 2021-01-20 main.tex
in IX again

A \bfit{fograph homomorphism} $\phi=\tuple{\phi,\substof\phi}$ is a
pair where $\phi\colon\gG\to\gH$ is a graph homomorphism between the
underlying graphs, and $\substof\phi$ is a variable renaming such that
for all $v\in\vG$ we have
$\labelof{\phi(v)}=\substof\phi(\labelof v)$. Note that this entails that
$\phi$ maps binders to binders and literals to literals. We say that a
fograph homomorphism $\phi\colon\gG \rightarrow \gH$ is
%%\bfit{label-preserving} if for all $v \in \vG, label(v) = label(f(v))$, and is
\bfit{existential-preserving} if for all existential binders $b$ in
$\gG$, the vertex $\phi(b)$ is an existential binder in $\gH$.

@@

\begin{lemma}\label{lem:shallow}
  Let $\vlinf{\rr}{}{\Scons B}{\Scons A}$ be an inference rule in
  $\FOMLS$ other than $\aiD$ and $\forall$. Then the sequent $\sqn{\cneg A,B}$ is provable in $\FOMLL$.
\end{lemma}
\juihsuan{For this, we need the introduction rule of $\fff$ in $\FOMLL$
(consider the $\aiD$ rule). But if we add it, we'll need the
$\mathsf{ai}\downarrow$ rule to have Lemma \ref{lem:MLL1->MLS1}}
\lutz{shit. so, we need another way to prove lemma~\ref{lem:MLS1->MLL1}, I'd like to avoid $\fff$-intro.}
\juihsuan{This lemma holds for all the rules except $\aiD$ (we may need to
mention that $A \fequ B \Rightarrow \ \vdash_{\FOMLL} \dual{B}, A$). The inductive
step for $\aiD$ in the proof of Lemma \ref{lem:MLS1->MLL1} can be solved by
replacing $t$ with $a \vlor \dual{a}$ and the $\ttt$-rule with the $\ax$
 rule followed by the $\vlor$ rule}\lutz{yes! and similar for $\forall$. Can you do that?}
\begin{proof}
  This is a straightforward exercise that we leave to the reader.
\end{proof}

@@

\begin{lemma}\label{lem:MLS1->MLL1}
  If a formula $C$ is provable in $\FOMLS$ then $\sqn C$ is provable in  $\FOMLL$. \todo{check: do we need rectified here?}
\end{lemma}

@@

\begin{lemma}
  \label{lem:me}
  The rules $\me,\mfaD,\mexD$ are derivable in $\set{\wrD,\cD,\fequ}$. 
\end{lemma}

\begin{proof}
  \todo{}\lutz{also check for the need of$\fequ$}

@@

\begin{lemma}
  Let $A$ and $B$ be formulas. Then
  \begin{equation*}
    \vlderivation{
      \vlde{\set{\wrD,\cD,\fequ}}{}{B}{
        \vlhy{A}}}
    \qquad
    \iff
    \qquad
    \vlderivation{
      \vlde{\set{\wrD,\acD,\me,\mfaD,\mexD,\fequ}}{}{B}{
        \vlhy{A}}}
  \end{equation*}
\end{lemma}

@@

For a derivation $\Deri$ in
$\set{\wrD,\acD,\me,\mfaD,\mexD,\fequ}$, we can now construct the
\bfit{rectification} $\rectif\Deri$ by rectifying each line of $\Deri$, yielding
a derivation in
$\set{\wrD,\acD,\me,\rectif\mfaD,\rectif\mexD,\fequ}$.

For each instance $\vlinf{\rr}{}{P}{Q}$ of an inference rule in
$\set{\wrD,\acD,\me,\rectif\mfaD,\rectif\mexD,\fequ}$ we can
define the \bfit{induced map} $\mapof{\rr}\colon\vgraphof
Q\to\vgraphof P$ which acts as the identity for
$\rr\in\set{\me,\fequ}$ and as the canonical injection for
$\rr=\wrD$. For $\rr=\acD$ it maps the vertices
corresponding to the two atoms in the premise to vertex of the
contracted atom in the conclusion, and for
$\rr\in\set{\rectif\cfaD,\rectif\mfaD,\rectif\mexD}$ it maps the two vertices
corresponding to the quantifiers in the premise to the one in the
conclusion (as acts as the identity on all other vertices).
For a
derivation $\Deri$ in
$\set{\wrD,\acD,\rectif\cfaD,\me,\rectif\mfaD,\rectif\mexD,\fequ}$ we can then define
the \bfit{induced map} $\mapof\Deri$ as the composition of the induced
maps of the rule instances in $\Deri$.
\juihsuan{maybe mention at least that the induced
maps define graph homomorphisms. Do we need to talk about the contexts
$\Sconhole$ here (induced maps act clearly as the identity on contexts but we need them for the composition)?}

@@

We say that a derivation $\Deri$ is \bfit{sane} if for every line $Q$
in $\Deri$ we have that $\graphof D$ is a fograph (i.e., all binders
are legal). Clearly, every rectified derivation is sane, but not vice
versa, as we might have multiple occurrences of bound varibles in $Q$,
such that $\graphof D$ is still a fograph.

\begin{lemma}\label{lem:cw->skew}
  Let $\vlderivation{
    \vlde{\set{\wrD,\acD,\cfaD,\me,\rectif\mfaD,\rectif\mexD,\fequ}}{\Deri}{B}{
      \vlhy{A}}}$ be a sane derivation. Then the induced map
  $\mapof\Deri\colon\graphof{A}\to\graphof{B}$ is a skew bifibration.
\end{lemma}

@@

\begin{proof}[Proof of Lemma~\ref{lem:cw->skew}]
  First, observe that for every inference rule
  $\rr\in\set{\wrD,\acD,\cfaD,\me,\rectif\mfaD,\rectif\mexD,\fequ}$ 
  the induced map $\mapof{\rr}\colon\vgraphof Q\to\vgraphof P$ defines
  a fibration on the corresponding binding graphs. \juihsuan{we may need to have
some explication here.} Therefore, their
  composition $\mapof\Deri$ is a fibration.

  Now we construct for
  $\Deri$ its propositional encoding $\PE\Deri$ by
  translating every line into its propositional encoding.
  \juihsuan{maybe mention that an instance of one of the other rules can be translated
into an instance of the same rule. It's trivial but may be worth mentioning.}
  The instances
  of the rules $\rectif\mfaD$ and $\rectif\mexD$ are replaced in two
  steps by:
  \begin{equation*}
    \vlderivation{
      \vlin{\rectif\acD}{}{\Scons{x\vlor(\PE{(Ax)}\vlor\PE{(Bx)})}}{
        \vlin{\fequ}{}{\Scons{(y\vlor z)\vlor(\PE{(Ay)}\vlor\PE{(Bz)})}}{
          \vlhy{\Scons{(y\vlor\PE{(Ay)})\vlor (z\vlor\PE{(Bz)})}}}}}
  \end{equation*}
  and
  \begin{equation*}
    \vlderivation{
      \vlin{\rectif\acD}{}{\Scons{x\vlan(\PE{(Ax)}\vlor\PE{(Bx)})}}{
        \vlin{\me}{}{\Scons{(y\vlor z)\vlan(\PE{(Ay)}\vlor\PE{(Bz)})}}{
          \vlhy{\Scons{(y\vlan\PE{(Ay)})\vlor (z\vlan\PE{(Bz)})}}}}}
  \end{equation*}
  respectively, where $\rectif\acD$ is a $\acD$ that renames the
  variabes---the propositional variable, as well as the first-order variable of
  the same name---as everything is recified, there is no ambiguity
  here.  This gives us a derivation $\vlderivation{
    \vlde{\set{\wrD,\acD,\rectif\acD,\me,\PE\fequ}}{\PE\Deri}{\PE B}{
      \vlhy{\PE A}}}$ such that $\mapof{\PE\Deri}=\mapof\Deri$. It has
  been shown in~\cite{str:07:RTA} that $\mapof{\PE\Deri}$ is a skew
  fibration (see
  also~\cite{hughes:invar,str:ral:tableaux19,str:RR-9048}). Hence,
  $\mapof\Deri$ is a skew fibration. That $\mapof{\PE\Deri}$ preserves
  exponentials follows immediately from its construction.
\end{proof}

@@

\begin{lemma}\label{skew->cw}
  Let $\gA$ and $\gB$ be fographs, let $\phi\colon\gA\to\gB$ be a skew
  bifibration, and let $A$ and $B$ be formulas with $\graphof A=\gA$
  and $\graphof B=\gB$. Then there are derivations
  \begin{equation*}
    \vlderivation{
      \vlde{\set{\wrD,\acD,\me,\rectif\mfaD,\rectif\mexD,\fequ}}{\rectif\Deri}{B}{
        \vlhy{A}}}
    \quand
    \vlderivation{
      \vlde{\set{\wrD,\acD,\me,\mfaD,\mexD,\fequ}}{\Deri}{B}{
        \vlhy{A\substof\phi}}}
  \end{equation*}
  such that $\mapof{\rectif\Deri}=\phi$ and $\rectif{\Deri}$ is a
  rectification of $\Deri$, and $\substof\phi$ is the substitution
  induced by $\phi$.
\end{lemma}

@@

\begin{proof}[Proof of Lemma~\ref{skew->cw}]
  By Lemma~\ref{lem:PE} we have $\gA=\graphof{\PE A}$ and
  $\gB=\graphof{\PE B}$.  Let $\vB'\subseteq\vB$ be the image of
  $\phi$, and let $\gB_1$ be the subgraph of $\gB$ induced by
  $\vB'$. Hence, we have two maps $\phi''\colon\gA\to\gB_1$ being a
  surjection and $\phi'\colon \gB_1\to\gB$ being an injection that
  reflects edges. \juihsuan{what do you mean by "reflect edges"?}\lutz{edge downstairs implies edge upstairs} Both, $\phi'$ and $\phi''$ remain skew
  bifibrations. Let us first look at $\phi'$.  Let $\PE B_1$ be the
  propositional formula obtained from $\PE B$ by removing all atoms
  that are not represented by vertices in $\vB'$. Then $\graphof{\PEp
    B_1}=\gB_1$.  By \cite[Proposition~7.6.1]{str:07:RTA}, we have a
  derivation $\vlder{\set{\wrD,\fequ}}{\PE\Deri_1}{\PE B}{\PEp B_1}$.
  A subformula of $\PE B$ is called \emph{weak} if it has been
  introduced by a weakening. All subformulas of a weak formula are
  also weak. Two weak subformulas $B'$ and $B''$ of $\PE B$ form a
  \emph{weak pair} if $B\fequ\Scons{B'\vlor B''}$ for some context
  $\Sconhole$. We now can assume without loss of generality that
  whenever weak subformulas $B'$ and $B''$ for a weak pair, they have
  been introduced by the same instance of $\wrD$ in
  $\PE\Deri_1$.\footnote{If $\PE\Deri_1$ is not of that shape, it can
  brought into this form by simple rule permutations, and then
  collapsing several weakenings into a single one.}  Now we show that
  $\PE\Deri_1$ can be lifted. For this, observe that whenever a
  weakening in $\PE\Deri_1$ deletes an atom $x\in\VAR$, it must also
  delete all atoms in the scope of the corresponding quantifier,
  because $\phi$ \juihsuan{should be $\phi'$ here} is a fibration on the binding graph. Hence, each line
  in $\PE\Deri_1$ is the propositional encoding $\PE P$ of a
  first-order formula $P$.  We now have to show
  that each instance of $\wrD$ is indeed a correct
  application in first-order logic. For this we have to inspect the
  cases a weakening happens inside a subformula $x\vlor C$ or $x\vlan
  C$ in $\PE\Deri_1$. There are the following cases:
  \begin{equation*}%\small
    \vlinf{}{}{S\cons{x\vlor D\vlor C}}{S\cons{x\vlor C}}
    \qquad
    \vlinf{}{}{S\cons{x\vlan (D\vlor C)}}{S\cons{x\vlan C}}
    \qquad
    \vlinf{}{}{S\cons{(x\vlor D)\vlan C}}{S\cons{x\vlan C}}
  In the first case the weakening happens inside the scope of a
  $\forall$-quantifier, and in the second case inside the scope of a
  $\exists$-quantifier. Both cases are unproblematic in first-order
  logic. However, in the third case an $\exists$-quantifier would be
  transformed into an $\forall$-quantifier. But as $\phi$ has to
  preserve existentials, this third case cannot occur.
  Thus we have a first order derivation
  $\vlder{\set{\wrD,\fequ}}{\Deri_1}{B}{B_1}$ with $\PE B_1=\PEp B_1$.

@@

  encoding. There are two reasons for this: (i) we might have $P\PE\fequ
  Q$ where $P$ is a propositional encoding but $Q$ is not, and (ii)
  the rule $\acD$ can duplicate an atom with name $x\in\VAR$. Let us write $\acDx$ for such instances. To
  address (i), we consider here formulas equivalent modulo $\fequ$,
  always knowing that we can add instances of $\fequ$ as needed. \juihsuan{this
does not seem clear to me. What if from $\PE A_1$ to $\PE B_1$ there are just a
bunch of $\PE\fequ$? What do we do in this case?} To

@@

  instances of $\acDx$ up until
  they either reach the top of the derivation or an instance of $\me$
  which separates the two atoms in the premise. More precisely, we
  consider the following inferene rule

@@

  analysis on the topmost rule that is not a $\fequ$.

@@

    propositional encoding. Hence, $S_1\conhole=\conhole\vlor E$ and
    $S_2\conhole=\conhole\vlor F$ and both $x$ are universals, and
    $E\vlor F$ contains all occurrences of $x$ bound by that

@@

        \vlinf{\acDeq}{}{\Scons{x\vlor (E\vlor F)}}{S\cons{(x\vlor E)\vlor (x\vlor F)}}

@@

    \juihsuan{We should say that $\Sconhole, E, $ and $F$ are also propositional
encodings. Here I think you kind of identify a formula with its prop. encoding.}
    \item $E$ is empty and $F$ is non-empty: We have

@@

        \vlinf{\acDeq}{}{\Scons{x\vlor F)}}{S\cons{x\vlor (x\vlor F)}}

@@

    \item $E$ is non-empty and $F$ is empty: This is similar to the previous case.
    \item $E$ and $F$ are both empty: This is impossible as the

>>>>

A \bfit{fograph homomorphism} $\phi=\tuple{\phi,\substof\phi}$ is a
pair where $\phi\colon\gG\to\gH$ is a graph homomorphism between the
underlying graphs, and $\substof\phi$, also called the \bfit{substitution induced by $\phi$} is a variable renaming such that
for all $v\in\vG$ we have
$\labelof{\phi(v)}=\substof\phi(\labelof v)$. Note that this entails that
$\phi$ maps binders to binders and literals to literals.
We say that a
fograph homomorphism $\phi\colon\gG \rightarrow \gH$ is
%%\bfit{label-preserving} if for all $v \in \vG, label(v) = label(f(v))$, and is
\bfit{existential-preserving} if for all existential binders $b$ in
$\gG$, the vertex $\phi(b)$ is an existential binder in $\gH$.

@@

\begin{lemma}\label{lem:shallow}
  Let $\vlinf{\rr}{}{\Scons B}{\Scons A}$ be an inference rule in
  $\FOMLS$ other than $\aiD$ and $\forall$. Then the sequent $\sqn{\cneg A,B}$ is provable in $\FOMLL$.
\end{lemma}
\juihsuan{For this, we need the introduction rule of $\fff$ in $\FOMLL$
(consider the $\aiD$ rule). But if we add it, we'll need the
$\mathsf{ai}\downarrow$ rule to have Lemma \ref{lem:MLL1->MLS1}}
\lutz{shit. so, we need another way to prove lemma~\ref{lem:MLS1->MLL1}, I'd like to avoid $\fff$-intro.}
\juihsuan{This lemma holds for all the rules except $\aiD$ (we may need to
mention that $A \fequ B \Rightarrow \ \vdash_{\FOMLL} \dual{B}, A$). The inductive
step for $\aiD$ in the proof of Lemma \ref{lem:MLS1->MLL1} can be solved by
replacing $t$ with $a \vlor \dual{a}$ and the $\ttt$-rule with the $\ax$
rule followed by the $\vlor$ rule}\lutz{yes! and similar for $\forall$. Can you do that?}

\lutz{actually NO! we can prove $\sqn{\fff,\ttt}$ using $\ax$ without a special rule for $\fff$. So can you please undo the changes and delete the comments?}
\begin{proof}
  This is a straightforward exercise that we leave to the reader. (Note that the $\ax$-rule can be applied to $\sqn{\fff,\ttt}$ in the cases of $\rr \in \set{\aiD, \faD}$.)
\end{proof}

@@

\begin{lemma}\label{lem:MLS1->MLL1}
  If a formula $C$ is provable in $\FOMLS$ then $\sqn C$ is provable in  $\FOMLL$. 
\end{lemma}

@@

\begin{lemma}
  \label{lem:me}
  $\cfaD,\me,\mfaD,\mexD$ are derivable in \hbox{$\set{\wrD,\cD,\fequ}$}. 
\end{lemma}

\begin{proof}
  \todo{}

@@

\begin{lemma}
  Let $A$ and $B$ be formulas. Then
  \begin{equation*}
    \vlderivation{
      \vlde{\set{\wrD,\cD,\fequ}}{}{B}{
        \vlhy{A}}}
    \qquad
    \iff
    \qquad
    \vlderivation{
      \vlde{\set{\wrD,\acD,\cfaD,\me,\mfaD,\mexD,\fequ}}{}{B}{
        \vlhy{A}}}
  \end{equation*}
\end{lemma}

@@

For a derivation $\Deri$ in
$\set{\wrD,\acD,\cfaD,\me,\mfaD,\mexD,\fequ}$, we can now construct the
\bfit{rectification} $\rectif\Deri$ by rectifying each line of $\Deri$, yielding
a derivation in
$\set{\wrD,\acD,\me,\rectif\mfaD,\rectif\mexD,\fequ}$.

For each instance $\vlinf{\rr}{}{P}{Q}$ of an inference rule in
$\set{\wrD,\acD,\rectif\cfaD,\me,\rectif\mfaD,\rectif\mexD,\fequ}$ we can
define the \bfit{induced map} $\mapof{\rr}\colon\vgraphof
Q\to\vgraphof P$ which acts as the identity for
$\rr\in\set{\me,\fequ}$ and as the canonical injection for
$\rr=\wrD$. For $\rr=\acD$ it maps the vertices
corresponding to the two atoms in the premise to vertex of the
contracted atom in the conclusion, and for
$\rr\in\set{\rectif\cfaD,\rectif\mfaD,\rectif\mexD}$ it maps the two vertices
corresponding to the quantifiers in the premise to the one in the
conclusion (as acts as the identity on all other vertices).
For a
derivation $\Deri$ in
$\set{\wrD,\acD,\rectif\cfaD,\me,\rectif\mfaD,\rectif\mexD,\fequ}$ we can then define
the \bfit{induced map} $\mapof\Deri$ as the composition of the induced
maps of the rule instances in $\Deri$.
\juihsuan{maybe mention at least that the induced
maps define graph homomorphisms. Do we need to talk about the contexts
$\Sconhole$ here (induced maps act clearly as the identity on contexts but we need them for the composition)?}
\lutz{For the context, I aleady say it is the identity. For the homom, it comes later}

@@

We say that a derivation $\Deri$ is \bfit{sane} if for every line $Q$
in $\Deri$ we have that $\graphof D$ is a fograph (i.e., all binders
are legal). Clearly, every rectified derivation is sane, but not vice
versa, as we might have multiple occurrences of bound variables in $Q$,
such that $\graphof Q$ is still a fograph.

\begin{lemma}\label{lem:cw->skew}
  Let $\vlderivation{
    \vlde{\set{\wrD,\acD,\rectif\cfaD,\me,\rectif\mfaD,\rectif\mexD,\fequ}}{\Deri}{B}{
      \vlhy{A}}}$ be a sane derivation. Then the induced map
  $\mapof\Deri\colon\graphof{A}\to\graphof{B}$ is a skew bifibration.
\end{lemma}

@@

\begin{proof}[Proof of Lemma~\ref{lem:cw->skew}]
  First, observe that for every inference rule
  $\rr\in\set{\wrD,\acD,\rectif\cfaD,\me,\rectif\mfaD,\rectif\mexD,\fequ}$ 
  the induced map $\mapof{\rr}\colon\vgraphof Q\to\vgraphof P$ defines a existential preserving graph homomorphism $\graphof Q\to\graphof P$ and  a fibration on the corresponding binding graphs. \juihsuan{we may need to have
some explication here.}\lutz{no} Therefore, their
  composition $\mapof\Deri$ has the same properties fibration.

  For showing that it is also a skew fibration, we construct for
  $\Deri$ its propositional encoding $\PE\Deri$ by
  translating every line into its propositional encoding.
  \juihsuan{maybe mention that an instance of one of the other rules can be translated
into an instance of the same rule. It's trivial but may be worth mentioning.}\lutz{done below}
  The instances
  of the rules $\rectif\mfaD$ and $\rectif\mexD$ are replaced in two
  steps by:
  \begin{equation*}
    \vlderivation{
      \vlin{\rectif\acD}{}{\Scons{x\vlor(\PE{(Ax)}\vlor\PE{(Bx)})}}{
        \vlin{\fequ}{}{\Scons{(y\vlor z)\vlor(\PE{(Ay)}\vlor\PE{(Bz)})}}{
          \vlhy{\Scons{(y\vlor\PE{(Ay)})\vlor (z\vlor\PE{(Bz)})}}}}}
  \end{equation*}
  and
  \begin{equation*}
    \vlderivation{
      \vlin{\rectif\acD}{}{\Scons{x\vlan(\PE{(Ax)}\vlor\PE{(Bx)})}}{
        \vlin{\me}{}{\Scons{(y\vlor z)\vlan(\PE{(Ay)}\vlor\PE{(Bz)})}}{
          \vlhy{\Scons{(y\vlan\PE{(Ay)})\vlor (z\vlan\PE{(Bz)})}}}}}
  \end{equation*}
  respectively, where $\rectif\acD$ is a $\acD$ that renames the
  variabes---the propositional variable, as well as the first-order
  variable of the same name---as everything is recified, there is no
  ambiguity here. Any instance of a rule $\wD$, $\acD$, $\me$, or
  $\fequ$ is translated to an instance of the same rule, and
  $\rectif\cfaD$ is translated to $\rectif\acD$.

  This gives us a derivation $\vlderivation{
    \vlde{\set{\wrD,\acD,\rectif\acD,\me,\PE\fequ}}{\PE\Deri}{\PE B}{
      \vlhy{\PE A}}}$ such that $\mapof{\PE\Deri}=\mapof\Deri$. It has
  been shown in~\cite{str:07:RTA} that $\mapof{\PE\Deri}$ is a skew
  fibration (see
  also~\cite{hughes:invar,str:ral:tableaux19,str:RR-9048}). Hence,
  $\mapof\Deri$ is a skew fibration.
  %%That $\mapof{\PE\Deri}$ preserves
  %%exponentials follows immediately from its construction.
\end{proof}

@@

\begin{lemma}\label{skew->cw}
  Let $\gA$ and $\gB$ be fographs, let $\phi\colon\gA\to\gB$ be a skew
  bifibration, and let $A$ and $B$ be formulas with $\graphof A=\gA$
  and $\graphof B=\gB$. Then there are derivations
  \begin{equation*}
    \vlderivation{
      \vlde{\set{\wrD,\acD,\rectif\cfaD,\me,\rectif\mfaD,\rectif\mexD,\fequ}}{\rectif\Deri}{B}{
        \vlhy{A}}}
    \quand
    \vlderivation{
      \vlde{\set{\wrD,\acD,\cfaD,\me,\mfaD,\mexD,\fequ}}{\Deri}{B}{
        \vlhy{A\substof\phi}}}
  \end{equation*}
  such that $\mapof{\rectif\Deri}=\phi$ and $\rectif{\Deri}$ is a
  rectification of $\Deri$, and $\substof\phi$ is the substitution
  induced by $\phi$.
\end{lemma}

@@

\begin{proof}[Proof of Lemma~\ref{skew->cw}]
  By Lemma~\ref{lem:PE} we have $\gA=\graphof{\PE A}$ and
  $\gB=\graphof{\PE B}$.  Let $\vB'\subseteq\vB$ be the image of
  $\phi$, and let $\gB_1$ be the subgraph of $\gB$ induced by
  $\vB'$. Hence, we have two maps $\phi''\colon\gA\to\gB_1$ being a
  surjection and $\phi'\colon \gB_1\to\gB$ being an injection that
  reflects edges. \juihsuan{what do you mean by "reflect edges"?}\lutz{edge downstairs implies edge upstairs} Both, $\phi'$ and $\phi''$ remain skew
  bifibrations. Let us first look at $\phi'$.  Let $\PEp B_1$ be the
  propositional formula obtained from $\PE B$ by removing all atoms
  that are not represented by vertices in $\vB'$. Then $\graphof{\PEp
    B_1}=\gB_1$.  By \cite[Proposition~7.6.1]{str:07:RTA}, we have a
  derivation $\vlder{\set{\wrD,\fequ}}{\PE\Deri_1}{\PE B}{\PEp B_1}$.
  A subformula of $\PE B$ is called \emph{weak} if it has been
  introduced by a weakening. All subformulas of a weak formula are
  also weak. Two weak subformulas $B'$ and $B''$ of $\PE B$ form a
  \emph{weak pair} if $\PE B\fequ\Scons{B'\vlor B''}$ for some context
  $\Sconhole$. We can assume without loss of generality that
  whenever weak subformulas $B'$ and $B''$ form a weak pair, they have
  been introduced by the same instance of $\wrD$ in
  $\PE\Deri_1$.\footnote{If $\PE\Deri_1$ is not of that shape, it can
  brought into this form by simple rule permutations, and then
  collapsing several weakenings into a single one.}  Now we show that
  $\PE\Deri_1$ can be lifted. For this, observe that whenever a
  weakening in $\PE\Deri_1$ deletes an atom $x\in\VAR$, it must also
  delete all atoms in the scope of the corresponding quantifier,
  because $\phi'$ is a fibration on the binding graph. Hence, each line
  in $\PE\Deri_1$ is the propositional encoding $\PE P$ of a
  first-order formula $P$.  We now have to show
  that each instance of $\wrD$ is indeed a correct
  application in first-order logic. For this we have to inspect the
  cases a weakening happens inside a subformula $x\vlor C$ or $x\vlan
  C$ in $\PE\Deri_1$. There are the following cases:
  \begin{equation*}%\small
    \vlinf{}{}{S\cons{x\vlor D\vlor C}}{S\cons{x\vlor C}}
    \qquad
    \vlinf{}{}{S\cons{x\vlan (D\vlor C)}}{S\cons{x\vlan C}}
    \qquad
    \vlinf{}{}{S\cons{(x\vlor D)\vlan C}}{S\cons{x\vlan C}}
  \end{equation*}
  In the first case the weakening happens inside the scope of a
  $\forall$-quantifier, and in the second case inside the scope of a
  $\exists$-quantifier. Both cases are unproblematic in first-order
  logic. However, in the third case an $\exists$-quantifier would be
  transformed into an $\forall$-quantifier. But as $\phi$ has to
  preserve existentials, this third case cannot occur.
  Thus we have a first order derivation
  $\vlder{\set{\wrD,\fequ}}{\Deri_1}{B}{B_1}$ with $\PE B_1=\PEp B_1$.

@@

  encoding. There are two reasons for this: (i) we might have
  $P\PE\fequ Q$ where $P$ is a propositional encoding but $Q$ is not,
  and (ii) the rule $\acD$ can duplicate an atom
  $x\in\VAR$. Let us write $\acDx$ for such instances. To address (i),
  we consider here formulas equivalent modulo $\fequ$, always knowing
  that we can add instances of $\fequ$ as needed.\footnote{Note that whenever we have formulas $P$ and $Q$ with $\PE P\PE\fequ\PE Q$ then $P\fequ Q$.} \juihsuan{this does
    not seem clear to me. What if from $\PE A_1$ to $\PE B_1$ there
    are just a bunch of $\PE\fequ$? What do we do in this case?}\lutz{see footnote} To

@@

  instances of $\acDx$ up until they either reach the top of the
  derivation or an instance of $\me$ which separates the two atoms in
  the premise. More precisely, we consider the following inferene rule

@@

  analysis on the topmost rule that is not a~$\fequ$.

@@

    propositional encoding. Hence, $S_1\conhole=\conhole\vlor \PE E$ and
    $S_2\conhole=\conhole\vlor \PE F$ and both $x$ are universals, and
    $\PE E\vlor \PE F$ contains all occurrences of $x$ bound by that

@@

        \vlinf{\acDeq}{}{\PE S\cons{x\vlor (\PE E\vlor \PE F)}}{\PE S\cons{(x\vlor \PE E)\vlor (x\vlor\PE F)}}

@@

      where $\PE S\conhole, \PE E, \PE F$ are the propositional encodings of $\Sconhole,E,F$, respectively.
    \item $\PE E$ is empty and $\PE F$ is non-empty: We have

@@

        \vlinf{\acDeq}{}{\PE S\cons{x\vlor \PE F)}}{\PE S\cons{x\vlor (x\vlor \PE F)}}

@@

    \item $\PE E$ is non-empty and $\PE F$ is empty: This is similar to the previous case.
    \item $\PE E$ and $\PE F$ are both empty: This is impossible as the
